\toolname: Property-Based Testing for MPC Security

### 2023-11-02 paper/main.tex
background & related work move

@@

Secure Multi-Party Computation (MPC or SMPC, depending on the text)
is a cryptographic process that allows some group of participating parties or machines
to compute a function that depends on all of their private inputs,
without any of them learning anything except the final result.
Implementations that work on arbitrary functions (representable as circuit computations in some field)
have existed since 1986~\cite{yao86}, % https://ieeexplore.ieee.org/document/4568207
but their use by the public is limited by their performance.
Some performance limitations are fundamental to the problem
(roughly, every branch of the computation must be executed so that the control-flow does not itself leak information),
but techniques have been found to improve performance somewhat.

The traditional expression of MPC assumes a pre-determined function F, "the functionality",
mapping the tuple of all participants' input vectors to the respective tuple of output vectors
(or to distributions of outputs, if F is non-deterministic).
In an \termOfArt{ideal world}, a trusted third party (the \termOfArt{ideal functionality}) would accept input messages
from all other parties, compute F, and then send the parties their respective outputs.
A \termOfArt{real world} protocol Π is said to \textit{correctly} implement F if, for all possible inputs,
Π results in all parties getting the same outputs as they would have in the ideal world.
The \textit{security} of Π is expressed in terms of a particular threat model and the participants' "\termOfArt{views}".

\paragraph{Passive Security}
supposes a computationally-bounded adversary
that can see everything happening inside a fixed subset of the participating machines.
This adversary cannot cause the parties to deviate from the protocol Π in any way;
we are interested in whether it can learn anything (beyond the outputs of the "corrupted" machines)
by observing the execution of Π.
As an example, this notion of security might be of interest if Alice trusted Bob to participate
in a distributed computation with her, but did not trust him to properly store or destroy the log-files
generated.
The \termOfArt{ideal world view}, $\mathsf{Ideal}_P(F,\vec{x})$, of a party P is what it sees when running F in the ideal world:
just its inputs $\vec{x}_P$ and its outputs $\vec{y}_P$.
The \termOfArt{real world view}, $\mathsf{Real}_P(Π, \vec{x})$, of a party P is everything it sees while executing its part of Π in the real world:
its inputs, outputs, all messages it receives, and all the bits of random tape it uses.
Both these concepts extend naturally to lists of parties.

\begin{definition}[Passive MPC Security]\label{def:passive-mpc}
    A protocol Π,
    involving participating parties P,
    that correctly implements functionality F,
    is secure against a passive adversary controlling C$\subset$P iff
    there exists a poly-time-computable function $\mathtt{SIM}$ s.t.
    for all combinations of inputs $\vec{x}$
    $$\begin{aligned}
        \mathsf{Real}_C(Π, \vec{x}) &\cong \mathtt{SIM}(\mathsf{Ideal}_C(F,\vec{x}))
    \end{aligned}$$
    where $\cong$ denotes distributional equivelence or indistinguishabilty.
\end{definition}

@@

\paragraph{Active Security}
(sometimes "malicious security")
supposes a computationally bounded adversary that can cause corrupted machines to take arbitrary actions:
breaking the choreography,
forging messages (to such extent as that's actually possible),
lying about computation results,
and/or playing along with the protocol for any period of time.
In contrast to passive adversaries, which are weaker than the threat models most real programmers worry about,
an active adversary subsumes any threat model (including adaptive corruption, depending how it's framed).
The corresponding security definition is a bit more involved, and we omit it for brevity.
Since we're not directly testing for active security in this work,
all else that needs to be said about it is that
any protocol which is not \termOfArt{passive secure} is automatically not \termOfArt{active secure}.

@@

\paragraph{Non-interference.}
While MPC can be expressed as a kind of non-interference property, this is only somewhat useful.
The non-interference can only be meant within the space of possible program traces \textit{posterior to the output}.
Trying to avoid such subtleties in the non-interference property by providing a correctly constrained \texttt{reveal} operator
isn't promising because constraints on \texttt{reveal} would need to recapitulate the original question of security
and examine the entire rest of the protocol as context.

TODO: maybe state it explicitly with the conditioning? Then we can say that no system can capture that property.

@@

Efficient inversion of programs and fragments of programs turns out to be important for determining if a protocol satisfies MPC security.
Since MPC protocols are probabilistic, what we really need is efficient posterior inference of program inputs in a
\termOfArt{probabilistic programming language} (PPL).

A PPL is a language that provides stochastic non-deterministic behavior as a language primitive
and provides meta-theory or auxiliary tooling to reason about the resulting distributions of program behavior.
In particular, PPLs commonly have semantics defined in terms of transformations on distributions,
and are designed to ensure that the resulting distributions are amenable to specified techniques for efficient
inference and sampling. Since even representing a fully-general multivariate distribution takes exponential space with respect to the number of variables,
PPLs typically restrict the programs and/or distributions they can handle.

For application to the subject of MPC protocols, a PPL would need to handle the two different kinds of uncertainty facing a hypothetical adversary:
1) Random values generated (by any party) at runtime have known distributions\footnote{
    \eg{} in python \texttt{secrets.SystemRandom().random()}
    will sample at runtime from the best floating-point approximation the machine can provide
    of the uniform distribution on $\mathbb{R}_{(0,1]}$
}.
2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
possible secrets representing their prior belief so that the secret inputs can also be modeled as random variables.

Hypothetically, supposing a protocol in a PPL providing a multivariate distribution over an appropriate subset of all runtime values,
if it were possible to efficiently infer and sample from the real-world views
(or other variables sufficient to deterministically compute a view)
\textit{posterior to} the corrupt outputs (and corrupt inputs),
then that would suffice as a simulator
\textit{if and only if} the inferred distribution were independent of the adversary's choice of prior distribution for the honest secrets.
So a direct test of MPC security (or a useful over-approximation) could be built from
A) a language meta-theory guarantee that such posterior inference and sampling would be efficient,
B) a test or guarantee of independence between the posterior view distribution and the exponentially-many parameters
of the prior distribution of honest secrets\footnote{
    Presumably such process would need to leave those parameters implicit.
    Attempting to represent the adversary's prior beliefs more compactly would not produce a valid proof or test of security.
}, and
C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs.
Such a system is not necessarily impossible, but we have so far failed to build it.

@@

\subsection{Formal-methods for cryptography}
Although a useful language-driven test for MPC security remains out of reach,
a variety of useful work aproaches the subject from different directions.

@@

Our python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
Annotation of which columns are available in the real and ideal worlds is provided in the header.
The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
and they're trained in parallel with \texttt{joblib.Parallel}.
The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.

@@

>>>>

\newcommand{\mynote}[2]
    {{\color{red} \fbox{\bfseries\sffamily\scriptsize#1}
    {\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}}~}
%\newcommand{\mynote}[2]{}
\newcommand{\todo}[1]{\mynote{TODO}{#1}}

@@

Secure Multi-Party Computation (MPC or SMPC, depending on the text)
is a cryptographic process that allows some group of participating parties or machines
to compute a function that depends on all of their private inputs,
without any of them learning anything except the final result.
Implementations that work on arbitrary functions (representable as circuit computations in some field)
have existed since 1986~\cite{yao1986generate}, % https://ieeexplore.ieee.org/document/4568207
but their use by the public is limited by their performance.
Some performance limitations are fundamental to the problem
(roughly, every branch of the computation must be executed so that the control-flow does not itself leak information),
but techniques have been found to improve performance somewhat.

The traditional expression of MPC assumes a pre-determined function F, the \emph{functionality},
mapping the tuple of all participants' input vectors to the respective tuple of output vectors
(or to distributions of outputs, if F is non-deterministic).
In an \termOfArt{ideal world}, a trusted third party (the \termOfArt{ideal functionality}) would accept input messages
from all other parties, compute F, and then send the parties their respective outputs.
A \termOfArt{real world} protocol Π is said to \textit{correctly} implement F if, for all possible inputs,
Π results in all parties getting the same outputs as they would have in the ideal world.
The \textit{security} of Π is expressed in terms of a particular threat model and the participants' "\termOfArt{views}".

\textbf{Passive security}
supposes a computationally-bounded adversary
that can see everything happening inside a fixed subset of the participating machines.
This adversary cannot cause the parties to deviate from the protocol Π in any way;
we are interested in whether it can learn anything (beyond the outputs of the "corrupted" machines)
by observing the execution of Π.
As an example, this notion of security might be of interest if Alice trusted Bob to participate
in a distributed computation with her, but did not trust him to properly store or destroy the log-files
generated.
The \termOfArt{ideal world view}, $\mathsf{Ideal}_P(F,\vec{x})$, of a party P is what it sees when running F in the ideal world:
just its inputs $\vec{x}_P$ and its outputs $\vec{y}_P$.
The \termOfArt{real world view}, $\mathsf{Real}_P(Π, \vec{x})$, of a party P is everything it sees while executing its part of Π in the real world:
its inputs, outputs, all messages it receives, and all the bits of random tape it uses.
Both these concepts extend naturally to lists of parties.

\todo{this is the first place the simulator is mentioned, but it's not described}
\begin{definition}[Passive MPC Security]\label{def:passive-mpc}
    A protocol Π,
    involving participating parties P,
    that correctly implements functionality F,
    is secure against a passive adversary controlling C$\subset$P iff
    there exists a poly-time-computable function $\mathtt{SIM}$ s.t.
    for all combinations of inputs $\vec{x}$
    $$\begin{aligned}
        \mathsf{Real}_C(Π, \vec{x}) &\cong \mathtt{SIM}(\mathsf{Ideal}_C(F,\vec{x}))
    \end{aligned}$$
    where $\cong$ denotes distributional equivelence or indistinguishabilty.
\end{definition}

@@

\textbf{Active security}
(sometimes "malicious security")
supposes a computationally bounded adversary that can cause corrupted machines to take arbitrary actions:
breaking the choreography,
forging messages,
lying about computation results,
and/or playing along with the protocol for any period of time.
In contrast to passive adversaries, which are weaker than the threat models most real programmers worry about,
an active adversary subsumes any threat model (including adaptive corruption, depending how it's framed).
The corresponding security definition is a bit more involved, and we omit it for brevity.
Since we're not directly testing for active security in this work,
all else that needs to be said about it is that
any protocol which is not \termOfArt{passive secure} is automatically not \termOfArt{active secure}.

@@

\paragraph{Non-interference.}
While MPC can be expressed as a kind of probabilistic non-interference property~\cite{haagh2018computer, almeida2018enforcing}, this is only somewhat useful.
The non-interference can only be meant within the space of possible program traces \textit{posterior to the output}.
Trying to avoid such subtleties in the non-interference property by providing a correctly constrained declassification operator
does not avoid this problem: constraints on declassification would need to recapitulate the original question of security
and examine the entire rest of the protocol as context.

@@

Since MPC security can be framed as a test of conditional independence, \termOfArt{probabilistic programming languages} (PPLs) with efficient inference and independence tests could be used to verify MPC security.
%
A PPL is a language that provides stochastic non-deterministic behavior as a language primitive
and provides meta-theory or auxiliary tooling to reason about the resulting distributions of program behavior.
In particular, PPLs commonly have semantics defined in terms of transformations on distributions,
and are designed to ensure that the resulting distributions are amenable to specified techniques for efficient
inference and sampling. Since even representing a fully-general multivariate distribution takes exponential space with respect to the number of variables,
PPLs typically restrict the programs and/or distributions they can handle.

For application to the subject of MPC protocols, a PPL would need to handle the two different kinds of uncertainty facing a hypothetical adversary:
(1) Random values generated (by any party) at runtime have known distributions\footnote{
    \eg{} in Python \texttt{secrets.SystemRandom().random()}
    will sample at runtime from the best floating-point approximation the machine can provide
    of the uniform distribution on $\mathbb{R}_{(0,1]}$
}, and
(2) The secret inputs of the honest parties are not stochastic at all, but the adversary can be asked to provide a distribution over
possible secrets representing their prior belief so that the secret inputs can also be modeled as random variables.

Hypothetically, supposing a protocol in a PPL providing a multivariate distribution over an appropriate subset of all runtime values,
if it were possible to efficiently infer and sample from the real-world views
(or other variables sufficient to deterministically compute a view)
\textit{posterior to} the corrupt outputs (and corrupt inputs),
then that would suffice as a simulator
\textit{if and only if} the inferred distribution were independent of the adversary's choice of prior distribution for the honest secrets.

A direct test of MPC security via a PPL would require \emph{both}
(A) a language meta-theory guarantee that such posterior inference and sampling would be efficient, \emph{and}
(B) a test or guarantee of independence between the posterior view distribution and the exponentially-many parameters
of the prior distribution of honest secrets\footnote{
    Presumably such process would need to leave those parameters implicit.
    Attempting to represent the adversary's prior beliefs more compactly would not produce a valid proof or test of security.
}, as well as
(C) the ability to quantify (A) and (B) over all possible corrupt secrets or public inputs, presumably by handling them symbolically.
While some PPLs can statically verify independence properties~\cite{gorinova2021conditional}, none can verify the kind of \emph{conditional} independence required for MPC security. Lilac~\cite{li2023lilac} provides a separation logic for reasoning about conditional independence, but is not automated.

@@

@@

Our Python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
Annotation of which columns are available in the real and ideal worlds is provided in the header.
The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
and they're trained in parallel with \texttt{joblib.Parallel}.
The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.

@@

\subsection{Formal-methods for cryptography}
Although automated language-based verification of MPC security is an open question,
a variety of useful work aproaches the subject from different directions.
